\chapter{Introduction}
\label{chap:introduction}

\section{Background}
\label{section:background}

The number of Thai stock investors has expanded substantially over the past 10 years.
Since 2015, the number of Thai stock investors has more than tripled, increasing from
approximately 1.2 million to around 4 million in 2024~\cite{SETStatistics2024}.
This piece of information highlights the growing interest in stock investments among Thai investors.

Despite the significant growth in the number of Thai stock investors, the Thai stock market remains highly volatile,
often trading in a sideways trend. This market behavior presents a challenging task to achieve exceptional returns,
requiring extensive research, analysis, and strategic decision-making.

Furthermore, many retail investors often make suboptimal investment decisions, caused by a combination of
overwhelming information, emotional biases, and market noise. As a result, there is a growing need for tools that can
improve their investment decisions and reduce the time and effort required.

\section{Problem Statement}
\label{section:problem-statement}

Analyzing stocks is a challenging task. Not only is it difficult to identify a good stock, but it also requires substantial time and effort
to read through reports and financial statements to identify a good company to invest in. Even professional investors with decades of experience
still take weeks or months to analyze large volumes of information for stock selection.

Apart from picking a good stock to invest in, investors often make emotion-based decisions, especially amateur and novice investors.
Suboptimal investment decisions increase the risk of capital loss. A good example would be buying stocks that are in a downtrend, thinking the
price should not go lower, or averaging down in an attempt to maintain their portfolios. However, these suboptimal decisions often lead to
further capital loss.

\section{Solution Overview}
\label{section:solution-overview}

Value Investing Intelligent Vanguard (VIIV) is a Thai-centric LLM-powered stock analyzer chatbot web application, designed to be an assistant for Thai stock investors,
aiding in stock analysis and data-driven investment decision-making. It gathers stock-related data from reliable and trusted sources to answer investors' questions with meaningful insights.

It addresses key pain points of Thai stock investors, including overwhelming information and time-consuming stock analysis. By answering stock-related questions,
VIIV eliminates the need for investors to read through hundred-page-long 56-1 One Reports for each company. It also helps identify stock price trends and professional analysts'
sentiments to assess investment opportunities and risks.

\subsection{Features}
\label{subsection:features}

\begin{enumerate}[leftmargin=80pt]
    \item \textbf{LLM-Powered Thai Stock Analyzer Chatbot}: A Thai language-centric large language model chatbot that provides information and insights to aid in stock analysis.
    \item \textbf{56-1 One Report and Financial Statement Analysis}: Analyzes yearly 56-1 One Reports and financial statements of companies to evaluate business models, executive strategies, and financial performance.
    \item \textbf{Stock Price Trend Identification with Technical Indicators}: Identifies stock price trends using technical indicators such as moving averages and moving average convergence divergence (MACD) to determine whether a stock is in an uptrend or downtrend.
    \item \textbf{IAA Consensus Sentiment Analysis}: Retrieves and analyzes the sentiment of the Investment Analysts Association (IAA) for each stock to assess opportunities and risks based on professional opinions.
\end{enumerate}

\section{Target User}
\label{section:target-user}

Our target users are \textbf{Thai stock investors}. This includes both novice and professional investors.
This group includes both male and female investors, with an age range of 25 to 50 years old.
Their education levels vary, but most demonstrate a strong ability to learn.

Since most of their investing and stock analysis require the use of computers and online tools,
Thai stock investors have a solid level of technical expertise, allowing them to utilize and adopt advanced investment software.

The main goal of Thai stock investors is to achieve high investment returns in the stock market while minimizing capital risk.
They regularly analyze individual stocks and market trends, seeking to invest in companies with high growth potential.
This introduces key challenges, including dealing with overwhelming information from large volumes of timely market data,
which may lead to missed opportunities and emotionally driven investment choices.

\section{Benefit}
\label{section:benefit}

Value Investing Intelligent Vanguard (VIIV) provides investors with various benefits, primarily by reducing the time and effort
required to analyze and study stocks. It allows users to ask stock-related questions in Thai and receive instant insights gathered
from complex financial data.

It helps assess opportunities and risks by identifying the trend of each individual stock using technical indicators
combining with the professional analysts' consensus. This ensures that investors have a clearer understanding of potential gains and
losses before making an investment decision.

These benefits enhance investment decision-making and improve investing performance in the stock market. They also help minimize the
risk of capital loss by informing potential opportunities and risks of a stock in both fundamental and technical aspects.

\section{Terminology}
\label{section:terminology}

\begin{itemize}
    \item \textbf{LLM (Large Language Model)}: A type of artificial intelligence model designed to understand and generate human-like text based on vast amounts of data.
    \item \textbf{56-1 One Report}: A financial document required by the Securities and Exchange Commission of Thailand (SEC) that provides a comprehensive overview of a company's financial performance, operations, and management.
    \item \textbf{MACD (Moving Average Convergence Divergence)}: A technical indicator used to analyze stock price trends by comparing short-term and long-term moving averages.
    \item \textbf{IAA (Investment Analysts Association)}: An organization that provides consensus reports and sentiment analysis from professional analysts regarding stock market conditions.
    \item \textbf{Fundamental Analysis}: A method of evaluating a stock by analyzing financial statements, management effectiveness, industry conditions, and other economic factors.
    \item \textbf{Technical Analysis}: A stock market analysis approach that uses historical price and volume data to identify trends and make investment decisions.
    \item \textbf{Downtrend}: A pattern in stock prices where the price consistently moves lower over a period of time.
    \item \textbf{Uptrend}: A pattern in stock prices where the price consistently moves higher over a period of time.
    \item \textbf{Sentiment Analysis}: A process of determining the general attitude or opinion of investors and analysts toward a particular stock or market trend.
\end{itemize}
